Introducing Distributed Dynamic Data-intensive (D3) Science: Understanding Applications and Infrastructure

### 2015-09-12 3dpas_intro.tex
finished refinement of introduction.

The focus of this paper is on a subspace of the large set of problems associated
with data-intensive sciences, namely that which is characterized by {\it both}
distributed and dynamic aspects. We introduce the concept of Dynamic Distributed
Data-intensive science and refer to this subspace of data-intensive science as
D3 science.  In order to appreciate the D3 subspace we will briefly discuss the
importance and role of distribution and the pervasiveness of dynamical aspects.

@@

Whether data are a priori distributed or localized, a critical
challenge is to find an efficient and effective means to either
distribute the initially localized data, or manage the distribution
(or localization) of a priori distributed data.

@@

@@

Analyzing the applications along an analytical space spanned by the
three ``D''s of data-intensive, distributed, and dynamic provides a
general and effective framework.  The diverse set of requirements to
support D3 scenarios points to the need for a careful understanding of
D3 characteristics as well as sophisticated, extensible, and
skillfully architected infrastructure and programming systems.

@@

Distribution can occur on many levels.
%, as shown in
%\figurename~\ref{fig:figures_terminology}.
In this paper, we use {\em locally distributed} to indicate that
data/compute resides on multiple nodes within one data center.
{\em Geographically distributed} describes the distribution of data and/or compute
across multiple data centers. Further, we define the degree as the
number of data sources and sites that are used for storing and
processing the data. A distributed application is an application that
is composed of entities (i.e., either storage/compute resource or a
data source) from at least two physical locations. In this work, our
focus is almost exclusively on geographically distributed data and
applications.

@@

\jhanote{refine next two paragraph...} Analogous to the distinction between programming in
the large versus system level programming, there is a distinction
between the ``cyberinfrastructure/systems considerations/design at the
large-scale'' versus ``CI considerations at the small-scale''. This
paper is primarily about CI/systems considerations at the large-scale.
Searching for common solutions and patterns at the large-scale is
likely to be more fruitful than at small-scales where customized tools
and optimization of low-level systems features is likely to return
greater yield.

Thus, subjecting a range of otherwise distinct applications to a
common analytical framework provides advantages to the infrastructure
scientists and cyberinfrastructure experts, namely suggestions for
where effort is both needed and likely to yield maximum returns. In
other words, as D3 applications evolve, our analysis can provide
guidance on where to focus effort on infrastructure.  Last but not
least, the advantages of providing a common but extensible vocabulary
that can be used by the entire community, application end-users,
developers and infrastructure scientists alike, in an emerging field
cannot be overstated. One may argue that such vocabulary and
terminology has been missing from the practice of distributed
cyberinfrastructure in general, with less than desirable consequences.

@@

Although the capacity to store data centrally is growing faster than Moore's
Law, it is not sufficient in all cases to be able to ``dump the data'' in one
place (or in the cloud.)  Oftentimes the costs (data transfer, storage,
consequences of catastrophic failure) prevent the centralization of data, and
decentralization---storing or place data in multiple locations---become prudent.

Conversely, not all production of data is centralized; in some cases there are
multiple producers (and consumers) of data, not to mention one-of-a-kind data
that must be shared across multiple consumers.  Furthermore in collaborative
science, data output from one analytical engine are often the input for the next
stage of the analytical pipeline.  Last but not least, distribution via
replication is a tested and effective approach to reliability and
fault-tolerance; we will argue that it is also potentially a great performance
enhancer and a means to ensure data consistency too.

@@

We would like the reader to keep in mind three primary context and
scope-setting points:

@@

\item Almost by definition, any attempt to capture the
  state-of-the-art of such a dynamic and emerging domain is bound to
  be fraught with limitations and issues of scoping.  A complete and
  rigorous survey of the landscape of BigData is bound to be obsolete
  by the time it emerges; so given the rate of change, we asked what
  could be done. What we present is a partial analysis motivated by a
  set of existing and active D3 applications, their trials and
  tribulations, successes and solutions. We believe there is merit in
  capturing the state-of-play as is, deriving a research agenda
  and set of specific yet crosscutting issues, and presenting these
  to the community.
\end{itemize}

@@

\subsection{Methodology}

@@

>>>>

@@

% Whether data are a priori distributed or localized, a critical
% challenge is to find an efficient and effective means to either
% distribute the initially localized data, or manage the distribution
% (or localization) of a priori distributed data.

@@

Subjecting a range of otherwise distinct applications to a
common analytical framework provides advantages to the infrastructure
scientists and cyberinfrastructure experts, namely suggestions for
where effort is both needed and likely to yield maximum returns. In
other words, as D3 applications evolve, our analysis can provide
guidance on where to focus effort on infrastructure.  Last but not
least, the advantages of providing a common but extensible vocabulary
that can be used by the entire community, application end-users,
developers and infrastructure scientists alike, in an emerging field
cannot be overstated. One may argue that such vocabulary and
terminology has been missing from the practice of distributed
cyberinfrastructure in general, with less than desirable consequences.

@@

The focus of this paper is on a subspace of the large set of problems associated
with data-intensive sciences, namely that which is characterized by {\it both}
distributed and dynamic aspects. We introduce the concept of Dynamic Distributed
Data-intensive science and refer to this subspace of data-intensive science as
D3 science.  Analyzing the applications along an analytical space spanned by the
three ``D''s of data-intensive, distributed, and dynamic provides a general and
effective framework.  The diverse set of requirements to support D3 scenarios
points to the need for a careful understanding of D3 characteristics as well as
sophisticated, extensible, and skillfully architected infrastructure and
programming systems.

%  In order to appreciate the D3 subspace we will briefly discuss the
% importance and role of distribution and the pervasiveness of dynamical aspects.

@@

%, as shown in  
%\figurename~\ref{fig:figures_terminology}.

@@

@@

\jhanote{refine next four paragraph...}

% Further, we define the degree as the number of data sources and sites that are
% used for storing and processing the data. A distributed application is an
% application that is composed of entities (i.e., either storage/compute resource
% or a data source) from at least two physical locations. In this work, our focus
% is almost exclusively on geographically distributed data and applications.

% Although the capacity to store data centrally is growing faster than Moore's
% Law, it is not sufficient in all cases to be able to ``dump the data'' in one
% place (or in the cloud.)  Oftentimes the costs (data transfer, storage,
% consequences of catastrophic failure) prevent the centralization of data, and
% decentralization---storing or place data in multiple locations---become prudent.

% Conversely, not all production of data is centralized; in some cases there are
% multiple producers (and consumers) of data, not to mention one-of-a-kind data
% that must be shared across multiple consumers.  Furthermore in collaborative
% science, data output from one analytical engine are often the input for the next
% stage of the analytical pipeline.  Last but not least, distribution via
% replication is a tested and effective approach to reliability and
% fault-tolerance; we will argue that it is also potentially a great performance
% enhancer and a means to ensure data consistency too.

@@

We would like the reader to keep in mind some context and scope-setting points:

@@

\item Almost by definition, any attempt to capture the state-of-the-art of such
  a dynamic and emerging domain is bound to be fraught with limitations and
  issues of scoping.  A complete and rigorous survey is bound to be obsolete by
  the time it emerges; so given the rate of change, we asked what could be
  done. What we present is a partial analysis motivated by a set of existing and
  active D3 applications, their trials and tribulations, successes and
  solutions. We believe there is merit in capturing the state-of-play as is,
  deriving a research agenda and set of specific yet crosscutting issues, and
  presenting these to the community.

\item Distribution can occur on many levels.  In this paper, we use {\em locally
    distributed} to indicate that data/compute resides on multiple nodes within
  one data center.  {\em Geographically distributed} describes the distribution
  of data and/or compute across multiple data centers.  Also, our focus is
  primarily on infrastructure considerations at the large-scale, as searching
  for common solutions and patterns at the large-scale is likely to be more
  fruitful than at small-scales where customized tools and optimization of
  low-level systems features is likely to return greater yield.

  % Analogous to the distinction between programming in the large versus system
  % level programming, there is a distinction between the ``infrastructure''
  % considerations at the large-scale versus the small-scale.

@@

@@

\section*{Methodology}
